---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
str(activity)
summary(activity)
```

## What is mean total number of steps taken per day?
```{r}
library(ggplot2)
totalSteps <- tapply(activity$steps, activity$date, FUN=sum, na.rm=TRUE)
qplot(totalSteps, geom = "histogram", xlab = "Total number of steps taken per day")
mean(totalSteps, na.rm = TRUE)
median(totalSteps, na.rm = TRUE)
```

## What is the average daily activity pattern?

```{r}

averages <- aggregate(x=list(steps=activity$steps), by=list(interval=activity$interval),
                      FUN=mean, na.rm=TRUE)
str(averages)
plot(averages$interval, averages$steps, type = "l", xlab = "5-minute interval",
     ylab = "Average number of steps taken")

averages <- data.frame(averages)
averages[which.max(averages$steps),]

```


## Imputing missing values


### The strategy was fill the missing values by the average for each 5-minute interval

```{r}
missing <- is.na(activity$steps)
table(missing)

fillMissing <- function(steps, interval){
    filled <- NA
    if(is.na(steps)){
        filled <- averages[averages$interval == interval, "steps"]
    }else{
        filled <- steps
    }
    return(filled)
}

fillActivity <- activity
fillActivity$steps <- mapply(fillMissing, fillActivity$steps, fillActivity$interval)

totalStepsFilled <- tapply(fillActivity$steps, fillActivity$date, FUN=sum, na.rm=TRUE)
qplot(totalStepsFilled, geom = "histogram", xlab = "Total number of steps taken per day")

## Show the mean and median
mean(totalStepsFilled, na.rm = TRUE)
median(totalStepsFilled, na.rm = TRUE)
```

Mean and median are higher after fill the missing values with the mean of each interval, because there are some days with steps values NA for any interval.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
fillActivity$date <- as.Date(fillActivity$date)
# create a vector of weekdays
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')

fillActivity$wDay <- factor((weekdays(fillActivity$date) %in% weekdays1), 
         levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))

str(fillActivity)

steps_by_interval <- aggregate(steps ~ interval + wDay, fillActivity, mean)

library(lattice)
xyplot(steps ~ interval | wDay, data = steps_by_interval, layout = c(1,2), type = "l",
       main="Average steps per day by interval")



```




